\chapter*{Acknowledgements}

I wish to say thanks to my supervisor, Professor Andrew Steane, who gave me a chance to study at one of the best places in the world for physics and guided me through these years. I am also extremely grateful for all fellow ion-trappers, past and present. To Professor Derek Stacey, whose logic was always there to help and never got tired of my questions; David Lucas, whose knowledge of every single knob in the lab and every variable in our software was invaluable, just as much as his enthusiasm; to Jonathan Home, Simon Webster, Matthew McDonnell, Nick Thomas and Jeff Sherman who showed me where I should be headed if I want to be a good researcher; to David Szwer, Spike Curtis, Alice Myerson and David Allcock for the long nights shared in the lab as well as the long chats every tea time, and for all the fun that is doing research in physics; to Graham Quelch and Rob Harris for their cheerful help in all things electronic and mechanical. I will certainly miss your company.

I must say thanks all my friends in St. Peter's College Middle Common Room and in many other colleges, who have shown me a completely different way one can be attached to people. Being Social Secretary for a year at St. Peter's gave me a chance to give back a little for all the things I got from you.

I want to say thanks to my parents, who brought me up to be a really curious person and encouraged me all the way: K\"osz\"on\"om!

And I owe a lot to YuJie, who supported me in the lows and shared the highs with me. Thank you.
